---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data

```{r}

actdata <- read.csv("activity.csv", colClasses = c("numeric", "character", "integer"))
actdata$date <- as.Date(actdata$date, "%Y-%m-%d")
head(actdata)

```



## What is mean total number of steps taken per day?

Histogram of total steps in each day:

```{r}
stepsbyday <- aggregate (x=actdata$steps, by=list(actdata$date), FUN=sum, na.rm=TRUE)
colnames(stepsbyday) = c("date", "totalsteps")
hist(stepsbyday$totalsteps, main = "Distribution of total steps by day", xlab = "total steps per day", breaks=10)
```

Mean and median of total steps per day:

```{r}
mean(stepsbyday$totalsteps)
```

```{r}
median(stepsbyday$totalsteps)
```


## What is the average daily activity pattern?

Time series plot:

```{r}
stepsbytime <- aggregate (x=actdata$steps, by=list(actdata$interval), FUN=mean, na.rm=TRUE)
colnames(stepsbytime) = c("time", "avgsteps")
plot(stepsbytime, type="l")
```

Time interval with highest average steps:

```{r}
stepsbytime [ which.max(stepsbytime$avgsteps), ]
```

## Imputing missing values

Total number of rows with NA:

```{r}
nrow (actdata[is.na(actdata), ])
```

Days and number of NA entries in these days:

```{r}
nadata = actdata[is.na(actdata),]
aggregate (x=nadata$interval, by=list(nadata$date), FUN=NROW)
```
There are 12*24=288 intervals per day. This means all entries in these days are NA


Imput missing value, using average steps of the same time interval:
```{r}

# save average steps at each time interval in a array using interval values as indexes
# plus 1 so that the first element is indexed as 1 instead of zero
avgstep_index <- 0
for (i in 1:nrow(stepsbytime)) {
  avgstep_index[ as.integer(1+stepsbytime$time[i]) ] <- stepsbytime$avgsteps[i]
}

#new data frame for storing data with imputed values
new_actdata = actdata

for (i in 1:nrow(new_actdata)) {
  if (is.na(new_actdata$steps[i])) {
    new_actdata$steps[i] = as.integer ( avgstep_index[ as.integer(new_actdata$interval[i]+1) ] )
  }
}

head(new_actdata)

```

ploting new histogram:

```{r}
new_stepsbyday <- aggregate (x=new_actdata$steps, by=list(new_actdata$date), FUN=sum)
colnames(new_stepsbyday) = c("date", "totalsteps")
hist(new_stepsbyday$totalsteps, main = "Distribution of total steps by day (with imputed values)", xlab = "total steps per day", breaks=10)
```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(lattice)

day <- weekdays(new_actdata$date)
daylevel <- vector()
for (i in 1:nrow(new_actdata)) {
if (day[i] == "Saturday") {
daylevel[i] <- "Weekend"
} else if (day[i] == "Sunday") {
daylevel[i] <- "Weekend"
} else {
daylevel[i] <- "Weekday"
}
}
new_actdata$daylevel <- daylevel
new_actdata$daylevel <- factor(new_actdata$daylevel)
new_stepsbytime <- aggregate(steps ~ interval + daylevel, data = new_actdata, mean)
colnames(new_stepsbytime) <- c("interval", "daylevel", "steps")

xyplot(steps ~ interval | daylevel, new_stepsbytime, type = "l", layout
= c(1, 2), xlab = "Interval", ylab = "Number of steps")

```

